Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
tweaks to abstract and intro

\begin{abstract}
  Verifying the correctness of numerical computations relies on first
  giving a specification of their behaviour. In many cases, a
  numerical computation is derived from an underlying mathematical
  model, but this is often difficult to use as a specification as it
  is highly decoupled from the implementation. We have been studying
  specification techniques in the middle ground between low-level and
  high-level specifications for numerical code. In this paper, we
  focus on \emph{stencil computations}---a common idiom in numerical
  code, but one which is error-prone due to fine-grained indexing
  errors. We hypothesise that in practise, stencil computations tend
  to have a regular shape which is amenable to static analysis and
  simple specification. We describe a abstract spatial specification
  language for stencil computations that can be embedded as
  annotations in source code, along with inference, checking, and
  specification and program synthesis procedures. We evaluate our
  language against a corpus of numerical Fortran code, and show that
  precise specifications can be given to roughly half of the
  identifiable stencil computations, and that the vast majority of
  stencil computations indeed have a simple, regular, static
  shape. This paper details our stencil specification language, its
  static semantics, inference, checking, and program synthesis
  procedures, our implementation, and our evaluation study.
\end{abstract}

@@

In this paper, we make the following contributions:
%
\begin{itemize}
\item we introduce our specification language for
stencil computations that captures many common forms
of data access pattern (Section~\ref{sec:lang})
along with a denotational model (Section~\ref{sec:semantics});

@@

\item we evaluate our approach, providing an 
implementation as an extension of CamFort,
an open-source program analysis and refactoring tool for Fortran
corpus. We use this implementation to evaluate our approach against
a corpus of scientific applications, totalling roughly one million
lines of code, and we study some verification case studies.
\dnote{insert results here}
\end{itemize}
%

@@

Our specification language is based on the observation
that most forms of array access in numerical code have
a fixed, statically-determined access pattern. For example, the
``\emph{five-point stencil}'' on a two-dimensional array reads from array
indices $(i, j)$, $(i-1, j)$, $(i+1, j)$, $(i, j-1)$, and $(i, j+1)$
for all $i, j$ within the inner boundary of the array (to avoid
out-of-bounds access at the edges). We revisit this hypothesis
in Section~\ref{sec:evaluation} where the inference of
such regular stencil patterns on a corpus of numerical programs (both
small and large). We found that indeed \dnote{..}.

>>>>

\begin{abstract}
  Verifying the correctness of numerical computations relies on first
  giving a specification of their behaviour. In many cases, a
  numerical computation is derived from an underlying mathematical
  model, but this is often difficult to use as a specification as it
  is highly decoupled from the implementation. We have been studying
  specification techniques in the middle ground between low-level and
  high-level specifications for numerical code. In this paper, we
  focus on \emph{stencil computations}---a common idiom in numerical
  code, but one which is error-prone due to fine-grained indexing
  errors. We hypothesise that in practise, stencil computations tend
  to have a regular shape which is amenable to static analysis and
  simple specification. We describe a abstract spatial specification
  language for stencil computations that can be embedded as
  annotations in source code, along with inference, checking, and
  specification and program synthesis procedures. We evaluate our
  language against a corpus of numerical Fortran code, and show that
  the vast majority of stencil computations indeed have a simple,
  regular, static shape. Around 10\% of the stencils found evoke a
  programming pattern for which indexing errors are plausible. We
  investigate two stencils in detail and find that we are able to
  detect the vast majority of potential errors. This paper details our
  stencil specification language, its static semantics, inference,
  checking, and program synthesis procedures, our implementation, and
  our evaluation study.
\end{abstract}

@@

In this paper, we make the following contributions:
%
\begin{itemize}
\item we introduce our specification language for
stencil computations that captures many common forms
of data access pattern (Section~\ref{sec:lang})
along with a sound denotational model (Section~\ref{sec:semantics});

@@

\item we evaluate our approach, providing an 
implementation as an extension of CamFort,
an open-source program analysis and refactoring tool for Fortran
corpus. We use this implementation to evaluate our approach against
a corpus of scientific applications, totalling roughly one million
lines of code, and we study some verification case studies;

\item we show that around 10\% of stencils are in a class that
  represent a plausible scenario for a programming error;

\item we consider two particular stencils in detail and show that we
  are able to detect the vast majority of (synthesised) possible
  indexing errors.
\end{itemize}
%

@@

Our specification language is based on the observation
that most forms of array access in numerical code have
a fixed, statically-determined access pattern. For example, the
``\emph{five-point stencil}'' on a two-dimensional array reads from array
indices $(i, j)$, $(i-1, j)$, $(i+1, j)$, $(i, j-1)$, and $(i, j+1)$
for all $i, j$ within the inner boundary of the array (to avoid
out-of-bounds access at the edges). We revisit this hypothesis
in Section~\ref{sec:evaluation} where the inference of
such regular stencil patterns on a corpus of numerical programs (both
small and large).
